Commutative Character Sheaves

### 2015-11-11 followup.tex
Outlines of the proofs for Section 3 are now complete.

\subsection{Commutative character sheaves on non-commutative groups}

@@

Now suppose $q : G \to G'$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G'
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $G'$.
Consider the functor
\[
q^* : \CS(G') \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(G') \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(G') \arrow{dl}{\text{comparison}}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

\begin{lemma}
If $q : G \to G'$ is an $H$-torsor in the fppf topology then the comparison functor $\CS(G') \to \CS_H(G)$ is an equivalence.
\end{lemma}

\begin{proof}
First we show that the comparison functor $\Loc(G') \to \Loc_H(G)$ is an equivalence. 
To see this, observe first that $\Loc$ is a stack on schemes over $\Fq$ in the Zariski topology.
Moreover, if $Y \to X$ is a flat surjective morphism of affine schemes over $\Fq$, then the comparison functor $\Loc(X) \to \Loc(Y\to X)$ is an equivalence, with $\Loc(Y\to X)$ defined as in [].
It follows that $\Loc$ is a stack over schemes over $\Fq$ in the fppf topology.
(This is a small variation on the proof that quasicoherent sheaves form a stack on sheaves in the fppf topology.)
Since $q : G\to G'$ is an $H$-torsor in the fppf topology, is now follows from descent theory that the comparison functor $\Loc(G') \to \Loc_H(G)$ is an equivalence.
The same argument shows that pull-back along $q\times q$ determines an equivalence $\Loc(G'\times G') \to \Loc_{H\times H}(G\times G)$.

@@

\subsection{Proofs}

@@

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$. 
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & \CCS(G\ab) \arrow{u} \arrow{dl} \\
\CCS(G) \arrow{u} \arrow{r}{\beta \mapsto \nu} & \arrow{u}  \CCS_{G\der}(G) & 
\end{tikzcd}
\]
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{G\der}(G) \to \CS'(G)$ is $\CCS(G)$.
To that end, suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ in Section~\ref{ssec:comcom}.
This determines a rigidification $r : \cs{L}_1 \to \EE$ by ....\todo{I've sorted this out, but have also run out of steam for the day. I'll write it up tomorrow.}
\end{proof}

%\begin{proof}[Proof of Proposition~\ref{prop:Gdiagram}]
%\end{proof}

>>>>

\subsection{Commutative character sheaves on non-commutative groups}\label{ssec:noncom}

@@

Now suppose $q : G \to Q$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
Q
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $Q$.
Consider the functor
\[
q^* : \CS(Q) \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(Q) \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(Q) \arrow{dl}{\text{comparison}}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

\begin{lemma}
If $q : G \to Q$ is an $H$-torsor in the fppf topology 
then the comparison functor $\CS(Q) \to \CS_H(G)$ is an equivalence.
\end{lemma}

\begin{proof}
First we recall that the comparison functor 
$\Loc(Q) \to \Loc_H(G)$ is an equivalence. 
To see this, observe first that $\Loc$ is a stack on schemes over $\Fq$ in the Zariski topology.
Moreover, if $Y \to X$ is a flat surjective morphism of affine schemes over $\Fq$, then the comparison functor $\Loc(X) \to \Loc(Y\to X)$ is an equivalence, with $\Loc(Y\to X)$ defined as in [].
It follows that $\Loc$ is a stack over schemes over $\Fq$ in the fppf topology.
(This is a small variation on the proof that quasicoherent sheaves form a stack on sheaves in the fppf topology.)
Since $q : G\to Q$ is an $H$-torsor in the fppf topology, is now follows from descent theory that the comparison functor $\Loc(Q) \to \Loc_H(G)$ is an equivalence.

The same argument shows that pull-back along $q\times q$ determines an equivalence $\Loc(Q\times Q) \to \Loc_{H\times H}(G\times G)$.

Now we claim that the comparison functor 
$\CS(Q)\to \CS_H(G)$ is also an equivalence.

@@

\begin{lemma}
If $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ may be apprehended as a morphism of $H\times H$-equivariant linear character sheaves on $G\times G$; likewise $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$. 
\end{lemma}

\begin{lemma}
If $(\cs{L},\mu,\beta)$ is a commutative character sheaf on $G$ then $\gamma : m^*\cs{L} \to \theta^* m^* \cs{L}$ is a morphism of commutative character sheaves on $G\times G$.
\end{lemma}

\subsection{Proofs for results from Section \ref{sec:defs}}

@@

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$. 
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{} \\
\arrow{u}{\text{forget}} \CS'(G)  \arrow{r}{\beta \mapsto \nu} &  \CS_{G\der}(G) \arrow{u}{\text{forget}}  & \CCS(G\ab) \arrow{u} \arrow{dl} \\
\CCS(G) \arrow{u} \arrow{r}{\beta \mapsto \nu} & \arrow{u}  \CCS_{G\der}(G) & 
\end{tikzcd}
\]
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{G\der}(G) \to \CS'(G)$ is $\CCS(G)$.
To that end, suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{ssec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as in Section~\ref{ssec:noncom}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}; 
in particular $\gamma = (q\times q)^* \xi$.

Using Lemma [], we may interpret the diagram in Definition~\ref{def:CCS} as a diagram in $\CS_{G\der\times G\der}(G\times G)$.
By Lemma [], this corresponds uniquely to a diagram in $\CS(G\ab\times G\ab)$, necessarily the diagram in Definition~\ref{def:CCScom}.
It now follows that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.
\end{proof}
